Supplementary Materials- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period

### 2021-02-04 analysis/paper/supplementary-material.Rmd
edit based on BF's comments, stop at model specification

<!--fit in a bigger historical picture in a pericolonial context: move to SI -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007]. In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007]. Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch. The Dutch asked local settlements to pay deer hides or rice as annual tributes but negotiating a delayed payment is allowed if it is needed [@Borao2009; @Kang2016]. In the annual event hold by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan stick or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements.

@@

A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994]. The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019]. For example, past trade can be conceptualized as a network of individual entities connected by shared similar objects or elements, such as the flow of goods, to represent their interactions [@Collar2015, pp. 4]. Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts. Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013]. This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.

@@

Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms. There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019]. Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019]. By quantifying those network properties, archaeologists can answer a wide range of research questions.

Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013]. Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials. Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010]. Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015]. Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].

@@

Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011]. For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby. Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics.

>>>>

<!--fit in a bigger historical picture in a pericolonial context: move to SI -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007]. In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007]. Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch. The Dutch asked local settlements to pay deer hides or rice as annual tributes but negotiating a delayed payment is allowed if it is needed [@Borao2009; @Kang2016]. In the annual event hold by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan stick or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements.

@@

A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994]. The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019]. For example, past trade can be conceptualized as a network of individual entities connected by shared similar objects or elements, such as the flow of goods, to represent their interactions [@Collar2015, pp. 4]. This concept can be applied to burial contexts where each burial is an actor linked by similar elements, such as burial goods. Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013]. This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.

@@

Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms. There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019]. Node level focuses on node properties in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019]. By quantifying those network properties, archaeologists can answer a wide range of research questions.

Examples include exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013]. Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials. Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010]. Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015]. Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].

@@

Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011]. For example, nodes with similar attributes are more likely to have a relationship, such as friendship between people with the same hobby. Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics.
